---
title: "What's Your Malfunction?"
author: "Victoria Zdanowicz"
date: "10/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

[1] Write a simple R function, Z.prop.test(), that can perform one- or two-sample Z-tests for proportion data, using the following guidelines:

Your function should take the following arguments: p1 and n1 (no default) representing the estimated proportion and sample size (i.e., based on your sample data); p2 and n2 (both defaulting to NULL) that contain a second sample’s proportion and sample size data in the event of a two-sample test; p0 (no default) as the expected value for the population proportion; and alternative (default “two.sided”) and conf.level (default 0.95), to be used in the same way as in the function t.test().

z = (observed statistic - expected statistic) / standard error

my.func <- function(arg1, arg2, arg3, ...) {
<commands>
return(output.object)
}

z.prop.test() <- function(p1, n1, p2=NULL, n2=NULL, p0, alternative = "two.sided", conf.level=0.95)
p1<- sum(v1)/n1
n1 <- length(v1)
p2 <- mean(v2)
n2 <- length(v2)
p0 = pi (expected proportion)

z1 <- (p1 - p0)/sqrt(1 - p0)/n1

pval <- 1 - pnorm(z1, lower.tail = TRUE) + pnorm(z1, lower.tail = FALSE)
pval

if(!is.numeric(c(a1, r, n))) return("The parameters must be numbers")

if p2 and/or n2 = NULL -> one-sample z test

z = (p-p0) / √p0(1-p0)/n

# great notes! I also found it helpful to see what variables I was dealing with before trying to create the function. 

```{r - tryinig to figure this out !!!!}
function (p1, n1, p2 = NULL, n2 = NULL, p0, alternative = "two-sided", conf.level = 0.95) 

  
  
  if (is.null(p2)|is.null(n2)) {
    Z <- (p1 - p0) / sqrt(p0 * (1-p0)/n1)
    PVAL <- 1 - pnorm(Z, lower.tail = TRUE) + pnorm(z, lower.tail = FALSE)
  
      lower <- p1 - qnorm(0.975) * sqrt(p1 * (1 - p1)/n1)
      upper <- p1 + qnorm(0.975) * sqrt(p1 * (1 - p1)/n1)
      ci <- c(lower, upper)
      one.sample.prop <- list(statistic = STATISTIC, parameter = PARAMETER, p.value = as.numeric(PVAL), 
                         conf.int = ci, alternative = alternative, method = METHOD)
      return(one.sample.prop)
      
      # Nice simple one tailed t-test, great job
  }
else {
  pstar <- (p1 + p2)/(n1 + n2)
  Z <- (p2 - p1)/sqrt((pstar * (1 - pstar)) * (1/length()))
  PVAL <- 1 - pnorm(Z, lower.tail = TRUE) + pnorm(Z, lower.tail = FALSE)
  crit <- qnorm(1 - alpha/2)  # identify critical values
  test <- p < -crit || p > crit  # boolean test
  test }

# Looks like there needed to be a bracket here
# good job with the two sample test!

p0 <- p0[OK]
if (any((p0 <= 0) | (p0 >= 1))) {
  stop("elements of 'p0' must be in (0,1)")

  PARAMETER <- k - 1 }

 else { # a little confused why these elses are turning up as errors. Curly brackets seem to be fine, but I am sure there is another reason 
  PARAMETER <- k
  names(NVAL) <- names(ESTIMATE)
}
names(PARAMETER) <- "df"
x <- cbind(x, n - x)
E <- cbind(n * p, n * (1 - p))
if (any(E < 5)) 
  warning("Chi-squared approximation may be incorrect")
STATISTIC <- sum((abs(x - E) - YATES)^2/E)
names(STATISTIC) <- "X-squared"
if (alternative == "two.sided") 
  PVAL <- pchisq(STATISTIC, PARAMETER, lower.tail = FALSE)
# 
else { # a little confused why these elses are turning up as errors. Curly brackets seem to be fine, but I am sure there is another reason 
  if (k == 1) 
    z <- sign(ESTIMATE - p) * sqrt(STATISTIC)
  else z <- sign(DELTA) * sqrt(STATISTIC)
  PVAL <- pnorm(z, lower.tail = (alternative == "less"))
# there was a bracket here that I think was causing problems 
RVAL <- list(statistic = STATISTIC, parameter = PARAMETER, 
             p.value = as.numeric(PVAL), estimate = ESTIMATE, null.value = NVAL, 
             conf.int = CINT, alternative = alternative, method = METHOD, 
             data.name = DNAME)
class(RVAL) <- "htest"
return(RVAL)
} 

# i've kind of lost you here with this section of code. I see you try to do the rule of thumb test first then I am confuused on what you are trying to do on the second part. 
```

# Just with a quick run through I am getting some errors. Let me see what I can try to improve here. 

# I am trying to fix the function and eliminate errors. 

# I was able to do a lot of the second question which I hope helped you when we talked about it in person. Looks like you got a lot further with how to create a function than I did though. Great attempt. I am glad we met in person to iron some of this out. Even if we don't get everything I think it is okay. I am sure most of next class will be on how to make functions considering everyone seems to be struggling with this. 


